---
title: "Scale space hydrograph analysis"
output: rmarkdown::html_vignette
bibliography: [references.bib, packages.bib]
biblio-style: apalike
link-citations: yes
vignette: >
  %\VignetteIndexEntry{scale-space}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## grwat tools for scale space transform

 __grwat__ package contains tools for performing scale space analysis of hydrograph. Initially developed by @witkin:1984, scale space transformation allows decomposition of a signal into hierarchically arranged multiscale components. You can use this tool to decompose, classify and compare hydrographs.

Currently __grwat__ depends on [Scale4C](https://bioconductor.org/packages/release/bioc/html/Scale4C.html) [@R-Scale4C] package from Biconductor.

__First__, use `grwat::ss_transform()` to get scale-space transformation for selected year and sigma (which is interpreted as maximum standard deviation in day units used for smoothing):

```{r}
library(grwat)

tab = read_separation('/Volumes/Work/_grwat/2018/out/1_Sever/Mezen_Malonisog/AllGrWat.txt')
(ssp = ss_transform(tab, year = 1978, sigma = 365))
```

__Second__, use `grwat::ss_tree()` to obtain a scale-space tree as a `tibble` object suited for visualization and analysis:
```{r}
(tree = ss_tree(ssp))
```

__Third__, use `grwat::plot_ss()` to plot resulting scale-space tree over the initial hydrograph:
```{r, fig.width=8, fig.height=4}
plot_ss(tree, tab, year = 1978)
```

## References