---
title: "Scale space hydrograph analysis"
output: rmarkdown::html_vignette
bibliography: [references.bib, packages.bib]
biblio-style: apalike
link-citations: yes
vignette: >
  %\VignetteIndexEntry{scale-space}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## grwat tools for scale space transform

 __grwat__ package contains tools for performing scale space analysis of hydrograph. Initially developed by @witkin:1984, scale space transformation allows decomposition of a signal into hierarchically arranged multiscale components. You can use this tool to decompose, classify and compare hydrographs.

Currently __grwat__ depends on [Scale4C](https://bioconductor.org/packages/release/bioc/html/Scale4C.html) [@R-Scale4C] package from Biconductor.

__First__, use `grwat::ss_transform()` to get scale-space transformation for selected year and sigma (which is interpreted as maximum standard deviation in day units used for smoothing):

```{r}
library(grwat)

tab = read_separation('/Volumes/Work/_grwat/2018/in/6_Terek/Baksan/AllGrWat.txt')
(sst = ss_transform(tab, year = 2016, sigma = 365))
```

__Second__, use `grwat::ss_tree()` to obtain a scale-space tree as a `tibble` object suited for visualization and analysis:
```{r}
(tree = ss_tree(sst))
```

__Third__, use `grwat::plot_ss()` to plot resulting scale-space tree over the initial hydrograph:
```{r, fig.width=8, fig.height=4}
plot_ss(tree, tab, year = 2016)
```

To _inverse_ the scale space tesselation on _Y_ axis, set `inverse = TRUE`:
```{r, fig.width=8, fig.height=4}
plot_ss(tree, tab, year = 2016, inverse = TRUE)
```

It is possible to plot scale space tesselation _without hydrograph_. Just omit `df` and `year` parameters in your call:
```{r, fig.width=8, fig.height=4}
plot_ss(tree)
```

You can also plot a __fingerprint map__ from which a scale space tesselation was traced back, using `Scale4C::plotTraceback()` (set `fileName = ""` to force plotting on screen device instead of a file):
```{r, fig.width=8, fig.height=6}
Scale4C::plotTraceback(sst, fileName = "")
```


## References